### 2021-12-15 docs/CV.Rmd
first commit

---
title: CV
author: Masatoshi Katabuchi
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
      fig_caption: yes
      theme: spacelab #readable #sandstone #spacelab #flatly
      highlight: pygments #tango #kate
      toc: TRUE
      toc_depth: 2
      number_sections: TRUE
      toc_float:
        collapsed: TRUE
        smooth_scroll: TRUE
        toc: true
---

```{r}
library(tidyverse)
```

# MVN version

## Model

First, we foucs on the species means.
Trait X (whole leaf LMA) for *i*th species is:
$$
X_i \sim N(\mu, \sigma_x^2).
$$

Trait Y (leaf disc LMA) for *i*th species is:

$$
Y_i = X_i + \epsilon_i
$$

where 
$$
\epsilon_i \sim N(0, \sigma_{\epsilon}^2) 
$$

**Q1: Do species mean of whole-leaf LMA and leaf disc LMA match well?**

This can be rephrased as testing the covariance between species mean of whole-leaf LMA and leaf disc LMA.
(i.e., 
$\mathrm{Cov}(X_i, Y_i)$
).
This will be simple and it depends on $\sigma_y$.

Then, let's look at the intraspecifc variation (ITV).
Trait x and y for *j*th leaf of *i*th species are assumed to follow a
multivariate normal distribution (MVN) because there should be positve covariances
between the trait values with species (i.e., individual that has large x should
have large y too).

$$
\begin{bmatrix}
x_{ij} \\
y_{ij} 
\end{bmatrix}
\sim MVN(
\begin{bmatrix}
\mu_{i} \\
\mu_{i} 
\end{bmatrix}
,
\begin{bmatrix}
\sigma_{xi}^2 & \rho\sigma_{xi}\sigma_{yi} \\
\rho\sigma_{xi}\sigma_{yi} & \sigma_{yi}^2 \\
\end{bmatrix}
)
$$

where $\mathrm{log}\sigma_{xi}$ and $\mathrm{log}\sigma_{yi}$ follow a MVN because there should be positive covariances
between the extents of intraspecifc variation (i.e., species that have large ITV
should have large ITV for both x and y traits).

$$
\begin{bmatrix}
\mathrm{log}\sigma_{x i} \\
\mathrm{log}\sigma_{y i}
\end{bmatrix}
\sim MVN(
\begin{bmatrix}
\tau_{x} \\
\tau_{y} 
\end{bmatrix}
,
\begin{bmatrix}
\hat{\sigma_{x}}^2 & \rho'\hat{\sigma_{x}}\hat{\sigma_{y}} \\
\rho'\hat{\sigma_{x}}\hat{\sigma_{y}} & \hat{\sigma_{y}}^2 \\
\end{bmatrix}
).
$$

Alternatively, ITV can be written as:

$$
x_{ij} \sim N(\mu_i, \sigma_{xi}^2)
$$

$$
y_{ij} = x_{ij} + \epsilon_{ij} 
$$

$$
\epsilon_{ij} \sim N(0,  \sigma_{\epsilon i}^2).
$$

CV of trait x for *i*th species is:

$$
CV_{xi} = \frac{\sigma_{xi}}{\mu_i}
$$

or

$$
CV = \sqrt{exp(\sigma_{xi}^2) - 1}.
$$

**Q2: Does the extent of intraspecifc variation match between the whole-leaf and
leaf disc estimates?**

This can be rephrased as testing the covariance between CV of LMA for each species estimated by whole-leaf and leaf disc.
(i.e., 
$\mathrm{Cov}(CV_{x}, CV_{y})$
).

## Example

```{r}
cv <- \(x) sd(x)/mean(x)
cv_ln <- \(x) sqrt(exp(var(x)) - 1)
```

```{r, eval=TRUE}
set.seed(123)
# n: sp
n <- 100
# m: trees for each species
m <- 10

mu <- 5
sig_x <- 1
sig_e <- 0.5
tau_x <- -1
sig_tx_hat <- 0.3
tau_y <- -1
sig_ty_hat <- 0.5

x <- rnorm(n, mu, sig_x)
y <- x + rnorm(n, 0, sig_e^2)
```

```{r}
rho0 <- 0.3
S0 <- c(sig_tx_hat^2, rho0 * sig_tx_hat * sig_ty_hat,
    rho0 * sig_tx_hat * sig_ty_hat, sig_ty_hat^2) |>
    matrix(ncol = 2)

sig_t <- mvtnorm::rmvnorm(n, mean = c(tau_x, tau_y), sigma = S0) |>
  exp()
```

```{r,eval=TRUE}
rho <- 0.5
tmp <- NULL
for (i in 1:length(mu_i)) {
  S <- c(sig_t[i,1]^2, rho * sig_t[i,1] * sig_t[i,2],
    rho * sig_t[i,1] * sig_t[i,2], sig_t[i,2]^2) |>
    matrix(ncol = 2)
  tmp <- rbind(tmp, mvtnorm::rmvnorm(m, mean = rep(mu_i[i], 2), sigma = S))
}
colnames(tmp) <- c("x", "y")

trait <- tmp |>
  as_tibble() |>
  mutate(sp = rep(1:n, each = m)) |>
  mutate(tree = rep(1:m, n))

cv_dat <- trait |>
  group_by(sp) |>
  summarise_at(
               .vars = vars(x, y),
               .funs = cv)

cv_dat2 <- trait |>
  group_by(sp) |>
  summarise_at(
               .vars = vars(x, y),
               .funs = cv_ln)
```

Species mean.

```{r}
plot(x, y)
cor.test(x, y)
```

Each tree.

```{r}
plot(y ~ x, trait)
cor.test(trait$y, trait$x)
```

CV.

```{r}
plot(y ~ x, cv_dat, log = "xy")
cor.test(log(cv_dat$y), log(cv_dat$x))
```

## Log-normal CV

$$
CV = \sqrt{exp(\mathrm{V}|x|) - 1}
$$

CV.

```{r}
plot(y ~ x, cv_dat2, log = "xy")
cor.test(log(cv_dat2$y), log(cv_dat2$x))
```

# Non-MVN version

## Example

```{r, eval=TRUE}
set.seed(123)
# n: sp
n <- 100
# m: trees for each species
m <- 10

mu <- 5
sig_x <- 1
sig_tx_hat <- 0.3
tau_x <- -1
mu_i <- rnorm(n, mu, sig_x)
sig_tx <- rnorm(n, tau_x, sig_tx_hat) |> exp()
x <- mapply(rnorm, m, mu_i, sig_tx)

tau_e <- -0.2
sig_te_hat <- 0.1
sig_te <- rnorm(n, tau_e, sig_te_hat) |> exp()
e <- mapply(rnorm, m, 0, sig_te^2)

y <- x + e

x <- as.numeric(x)
y <- as.numeric(y)

trait <- tibble(x, y) |>
  mutate(sp = rep(1:n, each = m)) |>
  mutate(tree = rep(1:m, n))

cv_dat <- trait |>
  group_by(sp) |>
  summarise_at(
               .vars = vars(x, y),
               .funs = cv)

mean_dat <- trait |>
  group_by(sp) |>
  summarise_at(
               .vars = vars(x, y),
               .funs = mean)

cv_dat2 <- trait |>
  group_by(sp) |>
  summarise_at(
               .vars = vars(x, y),
               .funs = cv_ln)

```

Species mean.

```{r}
plot(y ~ x, mean_dat)
cor.test(mean_dat$y, mean_dat$x)
```

Each tree.

```{r}
plot(y ~ x, trait)
cor.test(trait$y, trait$x)
```

CV.

```{r}
plot(y ~ x, cv_dat, log = "xy")
cor.test(log(cv_dat$y), log(cv_dat$x))
```

```{r, eval=FALSE, echo=FALSE}
x <- seq(0.1, 1.5, length = 80)
plot(x, sqrt(exp(x^2) - 1), type = "l")
lines(x, x, lty = 2)
lines(x, x^2, lty = 3)
lines(x, exp(x), lty = 4)
```