### 2022-04-26 ms/LMA_method.tex

%%
% Copyright (c) 2017 - 2020, Pascal Wagler;
% Copyright (c) 2014 - 2020, John MacFarlane
%
% All rights reserved.
%
% Redistribution and use in source and binary forms, with or without
% modification, are permitted provided that the following conditions
% are met:
%
% - Redistributions of source code must retain the above copyright
% notice, this list of conditions and the following disclaimer.
%
% - Redistributions in binary form must reproduce the above copyright
% notice, this list of conditions and the following disclaimer in the
% documentation and/or other materials provided with the distribution.
%
% - Neither the name of John MacFarlane nor the names of other
% contributors may be used to endorse or promote products derived
% from this software without specific prior written permission.
%
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
% "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
% LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS
% FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE
% COPYRIGHT OWNER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT,
% INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING,
% BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES;
% LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
% CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
% LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN
% ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
% POSSIBILITY OF SUCH DAMAGE.
%%

%%
% This is the Eisvogel pandoc LaTeX template.
%
% For usage information and examples visit the official GitHub page:
% https://github.com/Wandmalfarbe/pandoc-latex-template
%%

\DeclareUnicodeCharacter{2212}{-}

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*,table}{xcolor}
%
\documentclass[
  12pt,
  a4paper,
,tablecaptionabove
]{scrartcl}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1.2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\definecolor{default-linkcolor}{HTML}{A50000}
\definecolor{default-filecolor}{HTML}{A50000}
\definecolor{default-citecolor}{HTML}{4077C0}
\definecolor{default-urlcolor}{HTML}{4077C0}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  breaklinks=true,
  pdfcreator={LaTeX via pandoc with the Eisvogel template}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
% add backlinks to footnote references, cf. https://tex.stackexchange.com/questions/302266/make-footnote-clickable-both-ways
\usepackage{footnotebackref}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

% Make use of float-package and set default placement for figures to H.
% The option H means 'PUT IT HERE' (as  opposed to the standard h option which means 'You may put it here if you like').
\usepackage{float}
\floatplacement{figure}{H}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{lineno}
\linenumbers

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\date{}

%%
%% added
%%

%
% language specification
%
% If no language is specified, use English as the default main document language.
%

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
    % Workaround for bug in Polyglossia that breaks `\familydefault` when `\setmainlanguage` is used.
  % See https://github.com/Wandmalfarbe/pandoc-latex-template/issues/8
  % See https://github.com/reutenauer/polyglossia/issues/186
  % See https://github.com/reutenauer/polyglossia/issues/127
  \renewcommand*\familydefault{\sfdefault}
    % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

%
% for the background color of the title page
%

%
% break urls
%
\PassOptionsToPackage{hyphens}{url}

%
% When using babel or polyglossia with biblatex, loading csquotes is recommended
% to ensure that quoted texts are typeset according to the rules of your main language.
%
\usepackage{csquotes}

%
% captions
%
%\definecolor{caption-color}{HTML}{777777}
\definecolor{caption-color}{HTML}{37474F}
%\usepackage[font={stretch=1.2}, textfont={color=caption-color}, position=top, skip=4mm, labelfont=bf, singlelinecheck=false, justification=raggedright]{caption}
\usepackage[font={stretch=1}, textfont={color=caption-color}, position=top, skip=2mm, labelfont=bf, singlelinecheck=false, justification=raggedright]{caption}
\setcapindent{0em}

%
% blockquote
%
\definecolor{blockquote-border}{RGB}{221,221,221}
\definecolor{blockquote-text}{RGB}{119,119,119}
\usepackage{mdframed}
\newmdenv[rightline=false,bottomline=false,topline=false,linewidth=3pt,linecolor=blockquote-border,skipabove=\parskip]{customblockquote}
\renewenvironment{quote}{\begin{customblockquote}\list{}{\rightmargin=0em\leftmargin=0em}%
\item\relax\color{blockquote-text}\ignorespaces}{\unskip\unskip\endlist\end{customblockquote}}

%
% Source Sans Pro as the de­fault font fam­ily
% Source Code Pro for monospace text
%
% 'default' option sets the default
% font family to Source Sans Pro, not \sfdefault.
%
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
    \usepackage[default]{sourcesanspro}
  \usepackage{sourcecodepro}
  %\usepackage{}
  \else % if not pdftex
    \usepackage[default]{sourcesanspro}
  \usepackage{sourcecodepro}
  %\usepackage{}

  % XeLaTeX specific adjustments for straight quotes: https://tex.stackexchange.com/a/354887
  % This issue is already fixed (see https://github.com/silkeh/latex-sourcecodepro/pull/5) but the
  % fix is still unreleased.
  % TODO: Remove this workaround when the new version of sourcecodepro is released on CTAN.
  \ifxetex
    \makeatletter
    \defaultfontfeatures[\ttfamily]
      { Numbers   = \sourcecodepro@figurestyle,
        Scale     = \SourceCodePro@scale,
        Extension = .otf }
    \setmonofont
      [ UprightFont    = *-\sourcecodepro@regstyle,
        ItalicFont     = *-\sourcecodepro@regstyle It,
        BoldFont       = *-\sourcecodepro@boldstyle,
        BoldItalicFont = *-\sourcecodepro@boldstyle It ]
      {SourceCodePro}
    \makeatother
  \fi
  \fi

%
% heading color
%
\definecolor{heading-color}{RGB}{40,40,40}
\addtokomafont{section}{\color{heading-color}}
% When using the classes report, scrreprt, book,
% scrbook or memoir, uncomment the following line.
%\addtokomafont{chapter}{\color{heading-color}}

%
% variables for title and author
%
\usepackage{titling}
\title{}
\author{}

%
% tables
%

\definecolor{table-row-color}{HTML}{F5F5F5}
\definecolor{table-rule-color}{HTML}{999999}

%\arrayrulecolor{black!40}
\arrayrulecolor{table-rule-color}     % color of \toprule, \midrule, \bottomrule
\setlength\heavyrulewidth{0.3ex}      % thickness of \toprule, \bottomrule
\renewcommand{\arraystretch}{1.3}     % spacing (padding)

%
% remove paragraph indention
%
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines

%
%
% Listings
%
%

%
% header and footer
%
\usepackage{fancyhdr}

\fancypagestyle{eisvogel-header-footer}{
  \fancyhead{}
  \fancyfoot{}
  \lhead[]{}
  \chead[]{}
  \rhead[]{}
  %\lfoot[\thepage]{}
  \cfoot[]{}
  \cfoot[]{\thepage}
  \renewcommand{\headrulewidth}{0.0pt}
 % \renewcommand{\footrulewidth}{0.0pt}
 % \renewcommand{\headrulewidth}{0.4pt}
 % \renewcommand{\footrulewidth}{0.4pt}
}
\pagestyle{eisvogel-header-footer}

%%
%% end added
%%

\begin{document}

%%
%% begin titlepage
%%

%%
%% end titlepage
%%

Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

\[ \]

Phisamai Maenpuen\textsuperscript{1,2,3},
Masatoshi Katabuchi\textsuperscript{1*},
Yusuke Onoda\textsuperscript{4},
Cong Zhou\textsuperscript{1,2},
Jiao-Lin Zhang\textsuperscript{1,3*},
Ya-Jun Chen\textsuperscript{1,3,5}

\[ \]

\textsuperscript{1} CAS Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303, China

\textsuperscript{2} University of Chinese Academy of Sciences, Beijing 100049, China

\textsuperscript{3} Center of Plant Ecology, Core Botanical Gardens, Chinese Academy of Sciences, Yunnan 666303, China

\textsuperscript{4} Graduate School of Agriculture, Kyoto University, Kyoto 606-8502, Japan

\textsuperscript{5} Savanna Ecosystem Research Station, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Yuanjiang, Yunnan 6663300, China

\[ \]

\textbf{Corresponding Authors}:

Masatoshi Katabuchi

E-mail: \href{mailto:mattocci27@gmail.com}{\nolinkurl{mattocci27@gmail.com}}

Jiao-Lin Zhang

Tel: +86 691 871 3046;
Fax: +86 691 871 5070;
E-mail: \href{mailto:zjl@xtbg.org.cn}{\nolinkurl{zjl@xtbg.org.cn}}

\[ \]

Manuscript received \_\_\_\_\_\_\_; revision accepted \_\_\_\_\_\_\_.

\textbf{Running title}: Leaf disc and whole-leaf LMA

\newpage

\hypertarget{abstract}{%
\section{ABSTRACT}\label{abstract}}

\textbf{PREMISE:}
Leaf mass per area (LMA), which is the key trait in leaf economic spectrum and plant growth analysis, is measured from leaf discs or whole leaves.
These differences in the measurement methods may lead to large differences in the estimates of LMA values.

\textbf{METHODS:}
We determined LMA using leaf discs and whole-leaf (including petiole) for 334 woody species from a wide range of biomes (tropics, subtropics, savanna, and temperate) to examine what extent whole-leaf and leaf disc LMA match, whether the differences in the estimates are associated with leaf size and thickness, and whether intraspecifc variation for each species match.

\textbf{RESULTS:}
Disc-based estimates of species and individual mean LMA matched well whole-leaf estimates.
Thin and large leaves had weaker correlations between leaf disc and whole-leaf estimates.
Variability of LMA estimates within species (i.e., amount of intraspecific variation) only weakly matched between whole-leaf and disc-based estimates (\emph{R\textsuperscript{2}} = 0.33).
We also found that sample with small total dry mass of leaf discs tended to show greater differences between leaf disc and whole-leaf LMA.

\textbf{CONCLUSIONS:}
Mean values of leaf disc LMA are generally good proxies for mean values of whole-leaf LMA with appropriate calibration (9.4\% differences in our samples), but their accuracy depends on leaf size, leaf thickness, and sizes of leaf punch.
Therefore, using an appropriate size of leaf punch is important for obtaining stable estimates of leaf disc LMA that matches well with whole-leaf LMA.
Quantifying trait variation using leaf disc should be considered with caution.

\textbf{KEY WORDS:}
intraspecific variation,
leaf density,
leaf economic spectrum,
leaf heterogeneity,
leaf punch,
leaf size,
leaf thickness,
petiole,
specific leaf area,
within-leaf variation

\hypertarget{introduction}{%
\section{INTRODUCTION}\label{introduction}}

Leaf mass per area (LMA), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area thus reflecting a cost for light interception, which is the key trait of the leaf economic spectrum (LES) (\protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}; \protect\hyperlink{ref-Poorter2009}{Poorter et al., 2009}; \protect\hyperlink{ref-Onoda2017}{Onoda et al., 2017}).
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover (\protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}).
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns (\protect\hyperlink{ref-Garnier1994}{Garnier and Laurent, 1994}; \protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}; \protect\hyperlink{ref-Reich2014}{Reich, 2014}; \protect\hyperlink{ref-Diaz2016}{Díaz et al., 2016}).
The LMA is also frequently used in plant growth analysis (\protect\hyperlink{ref-Evans1972}{Evans, 1972}; \protect\hyperlink{ref-Poorter2014}{Poorter et al., 2014}; \protect\hyperlink{ref-Falster2018}{Falster et al., 2018}) because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR \(\times\) LMR \(\times\) LMA\textsuperscript{-1}).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species (i.e., `soft' trait, \protect\hyperlink{ref-Diaz2004}{Díaz et al., 2004}).
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago (\protect\hyperlink{ref-Hanson1917}{Hanson, 1917}).

LMA values can be determined by either
1) a whole leaf including or excluding petioles, and
2) a leaf disc excluding all major veins or petioles (\protect\hyperlink{ref-Perez-Harguindeguy2013}{Pérez-Harguindeguy et al., 2013}; \protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}).
Although some studies do not make clear what protocol was followed, these differences in the measurement methods may lead to large discrepancies in the estimates of LMA values because the major vein allocation (major vein, which includes first to second or third-order veins, volume per area) has been reported to be one of the main determinants for the variation in LMA (\protect\hyperlink{ref-John2017}{John et al., 2017}).
Since larger leaves tend to invest more of their mass into dense midribs and petiole for support (\protect\hyperlink{ref-Niinemets2006}{Niinemets et al., 2006}, \protect\hyperlink{ref-Niinemets2007}{2007}), and thinner leaves have clearly visible and large-diameter veins with less uniform leaf structure (i.e., kite-type leaves; (\protect\hyperlink{ref-Grubb1986}{Grubb, 1986})), discrepancies in the estimates of whole-leaf LMA and leaf disc LMA might be greater for larger and thinner leaves.
If leaf disc LMA largely and consistently underestimates whole-leaf LMA, some calibrations may be required to combine or compare those different estimates (\protect\hyperlink{ref-Kraft2008}{Kraft et al., 2008}; \protect\hyperlink{ref-Onoda2011}{Onoda et al., 2011}).
If divergencies between the estimates based on the different methods are large and inconsistent (i.e., low \emph{R\textsuperscript{2}} values), this is difficult to calibrate and should inflate errors in the subsequent analyses.
To date, however, only a database from a single region, Panama Plant Traits Database (\protect\hyperlink{ref-Wright2010}{Wright et al., 2010}), is available in the TRY (\protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}) that have both estimates of LMA from small leaf discs and whole leaves including petioles, limiting of our understanding of under what conditions estimates of LMA from leaf discs are reliable estimates of whole-leaf LMA.

Given that leaf disc samples have more sources of intraspecific trait variation (ITV) than whole-leaf samples, the effects of discrepancies between whole-leaf LMA and leaf disc LMA might be large when ITV is quantified as a coefficient of variance (CV).
Sources of ITV for whole-leaf samples is variation among individuals within the same species and variation among leaves within the same individuals (\protect\hyperlink{ref-Messier2010}{Messier et al., 2010}, \protect\hyperlink{ref-Messier2017}{2017}).
An additional source of trait variation for leaf disc samples is variation among leaf discs within the same leaves.
Thus, extents of ITV for each species based on whole-leaf and leaf discs samples may not match.
Despite the importance of ITV in community ecology (\protect\hyperlink{ref-Kichenin2013}{Kichenin et al., 2013}; \protect\hyperlink{ref-Siefert2015}{Siefert et al., 2015}), the effect of different measurement methods on the extent of ITV has been largely ignored.

In this study, we aimed to investigate what extent whole-leaf (including petiole) and leaf disc LMA match, whether the relationships between whole-leaf LMA and leaf disc LMA varied in leaf sizes and leaf thicknesses, and whether amounts of intraspecifc variations for each match between whole-leaf and leaf disc based estimates.
We collected a total of 334 woody species from four biomes (tropics, subtropics, savanna, and warm-temperate) to cover the wide range of geography.
We predicted that the overall relationship between whole-leaf LMA and disc LMA would be weaker than previous reported LMA data (\emph{R\textsuperscript{2}} = 0.92 for whole-leaves in Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} = 0.92 for leaf laminas in Onoda et al. (\protect\hyperlink{ref-Onoda2011}{2011})) because we collected data with various leaf sizes from a wide range of climates.
We also predicted that species with larger and/or thinner leaves would show weaker correlations.
Finally, we predicted that amounts of intraspecifc variations for each species do not match well between whole-leaf and leaf disc based estimates.

\hypertarget{materials-and-methods}{%
\section{MATERIALS AND METHODS}\label{materials-and-methods}}

\emph{Data sources}

We used newly compiled individual-level plant datasets form a wide ranges of biomes in China and Japan (Table S1) to examine the relationship between whole-leaf and leaf disc-based estimates of leaf traits.
First, the Yunnan dataset is from three forest plots in Yunnan province, Southwest China, which includes a tropical rain forest (TRF), a hot-dry savanna ecosystem (HDS), and a subtropical evergreen wet forest (STF), with a total of species and individuals.
Second, the Yakushima data is from a warm-temperate forest in Yakushima island, Japan, with a total of 193 species and individuals.
In total, we used 334 woody species and individuals.
The methodologies for trait sampling and measurement are slightly different between the Yunnan and the Yakushima dataset, which we describe bellow.

\emph{Measurements of leaf disc and whole-leaf LMA}

At the TRF site, we collected leaf samples from trees within reach of a canopy crane (88 m tall with a 60 m long boom).
In the HDS, and STF sites, we used a 12-m long pruner to collect samples for most of the target species.
In case of tall individuals that out of the reach of pruner, we used a rope climbing to reach the canopy, then sampled branches with the long pruner.
In Yakushima, we used 15-m poles for our leaf sampling.
At least six sun-exposed healthy leaves were sampled from each of 3 to 6 individuals for each species.
Trait values were averaged at the species- and individual-level for the analyses.
For compound-leaved species, we referred a leaflet (the minimum photosynthetic unit) as a single leaf in our study.

To determine what extent LMA values were affected by the different methods, LMA were determined (i) LMA based on a whole leaf including petioles and (ii) LMA based on leaf disc avoiding thick veins.
Three 0.6-cm diameter discs were taken from the base, middle and tip of each leaf using a hole punch in the Yunnan dataset.
Two 1.0-cm diameter discs were taken from each leaf were in the Yakushima dataset.
Fresh leaf area (LA; cm\textsuperscript{-2}) of leaf materials (whole leaves including petioles and midribs) was measured using a scanner and ImageJ software by the R package \emph{LeafArea} (\protect\hyperlink{ref-Katabuchi2015}{Katabuchi, 2015}).
Fresh leaf thickness (LT; mm) was measured using a micrometer in three points at the base, middle and tip of the leaf using a micrometer (Mitutoyo 293-240, Mitutoyo Corporation, Kawasaki, Japan) with a precision at 0.001 mm.
Thickness of each leaf disc was also measured for the Yunnan dataset but not for the Yakushima dataset, which enables us to estimate whole-leaf and leaf disc leaf tissue density (LD; g cm\textsuperscript{-3}) separately in the Yunnan dataset.
Leaf dry mass (whole leaves including petioles and midribs, and leaf discs) was recorded after drying 80°C for over 48 h to constant weight using an electronic scale (Mettler-Toledo MS204TS, Columbus, OH, U.S.A.) with a precision at 0.0001 g.
In the subset of Yakushima dataset, leaf lamina and petiole dry mass were also determined separately (86 out of 193 species) to quantify leaf support costs.
Dry mass for leaf discs were recorded 2-3 discs together for each leaf.
Based on those measurements of whole-leaf and leaf disc samples, LMA (g m\textsuperscript{-2}) was calculated as the ratio of leaf dry mass to fresh leaf area, and LD was calculated as the ratio of LMA to LT.
We do not present results of LD in the main text, because differences between whole-leaf LD and leaf disc LD only depends on the ratio between whole-leaf LMA and leaf disc LMA and measurement errors of leaf thickness (Appendix S2).

\emph{Effects of leaf sizes and thicknesses on the relationship between whole-leaf and leaf disc LMA}

To quantify the effect of leaf sizes and thicknesses on the relationship between whole-leaf and leaf disc LMA, we built hierarchical Bayes models.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of leaf size and leaf thickness on the log-scale;

\begin{align}
\mathrm{ln}(LMA_{Wi}) & \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i) \label{eq:LMA}\\
\tilde{\mu_i} & = \beta_0 + \beta_1 \mathrm{ln}(LA_{i}) + \beta_2 \mathrm{ln}(LT_{i}) \label{eq:mu}
\end{align}

where \(LMA_{Wi}\), \(LMA_{Di}\), \(LA_{i}\), \(LT_{i}\) are the whole-leaf LMA, leaf disc LMA, leaf area and leaf thickness for tree \emph{i}, respectively.
\(\tilde{\mu_i}\) is the effect of leaf area and leaf thickness on the mean estimates for the whole-leaf LMA of tree \emph{i}, and \(\sigma_i\) is the variance of whole-leaf LMA for tree \emph{i}, and \(\beta\) are coefficients.

The variance of whole-leaf LMA (\(\sigma_i\)) was assumed to follow a normal distribution;

\begin{align}
\mathrm{ln}(\sigma_i) & \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega) \label{eq:ln-sigma} \\
\tilde{\sigma_i} & = \gamma_0 + \gamma_1 \mathrm{ln}(LA_{i}) + \gamma_2 \mathrm{ln}(LT_{i}) \label{eq:tilde-sigma}
\end{align}

where \(\tilde{\sigma_i}\) is the effect of leaf area and leaf thickness on the variance of whole-leaf LMA,
\(\omega\) is a scaling hyperparameter, and \(\gamma\) are coefficients.
To allow comparisons among parameter estimates, \(\mathrm{ln}(LA)\) and \(\mathrm{ln}(LT)\) were scaled to a mean of 0 and standard deviation of 1.

Posterior distributions of all parameters were estimated using the Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan (\protect\hyperlink{ref-Carpenter2017}{Carpenter et al., 2017}) using the weakly-informative priors (\protect\hyperlink{ref-Gelman2008}{Gelman et al., 2008}).
See supplement XXXX for more detail.
The Stan code use to fit models is available from Github at: \url{https://github.com/mattocci27/leaf-disc}.
Convergence of the posterior distribution was assessed with the Gelman-Rubin statistic with a convergence threshold of 1.1 for all parameters (\protect\hyperlink{ref-Gelman2013}{Gelman et al., 2013}).

\emph{Bivariate trait relationships}

Bivariate trait relationships between disc-based and whole-leaf log-transformed LMA were summarized with variance explained (\emph{R\textsuperscript{2}}) based on Pearson correlations and with standardized major axis (SMA) regressions (\protect\hyperlink{ref-Warton2006}{Warton et al., 2006}) across all the samples and different leaf area and thickness groups.
Species and individual groups were divided into four categories based on the medians of leaf area and leaf thickness across all species and all individuals, respectively:
thin and small-leaved, thin and large-leaved, thick and small-leaved, thick and large-leaved species (individuals).
Median values are as following:
small-leaved species (LA \textless{} cm\textsuperscript{2}),
thin-leaved species (LT \textless{} mm),
small-leaved individuals (LA \textless{} cm\textsuperscript{2}),
and thin-leaved individuals (LT \textless{} mm).

\emph{Intraspecifc variation}

The amount of intraspecific variation (ITV) was calculated using Bao's CV estimator (\protect\hyperlink{ref-Yang2020}{Yang et al., 2020});

\begin{align}
CV = CV_1 - \frac{CV_1^3}{N} + \frac{CV_1}{4N} + \frac{CV_1^2 \gamma_1}{2N} + \frac{CV_1 \gamma_2}{8N} \label{eq:CV}
\end{align}

where \(CV_1 = s/\bar{x}\), \(s\) is the sample standard deviation and \(\bar{x}\) is the sample mean, \emph{N} is the sample size, and \(\gamma_1\) and \(\gamma_2\) are the skewness and kurtosis of the trait distribution.
Bao's CV estimator is a robust method to quantify ITV based on coefficient of variation (CV) when sample size is small (\protect\hyperlink{ref-Yang2020}{Yang et al., 2020}).

SMA analyses were run using the R package \emph{smatr} (\protect\hyperlink{ref-Warton2012a}{Warton et al., 2012}); this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 2000 bootstraps.
All statistical analyses were conducted in R version 4.1.3 (\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}).

\hypertarget{results}{%
\section{RESULTS}\label{results}}

Disc-based estimates of species mean LMA matched well whole-leaf estimates (Table 1; \emph{R\textsuperscript{2}} = for all species).
Species with thicker and smaller leaves had stronger correlations between disc-based and whole-leaf estimates of LMA (Table 1, Fig. 1).
There was a negative correlation between species mean leaf area and leaf thickness (\emph{r} = , \emph{P} , \emph{n} = ).
When we pooled all the species, there were significant differences in intercepts for species mean LMA (Table 1), suggesting that the whole-leaf estimates of LMA were 9.4\% greater than disc-based estimates in the arithmetic scale.
Large-leaved species also had significant positive intercepts, suggesting that whole-leaf estimates of LMA were 22.1\% greater than disc-based estimates in the arithmetic scale (Table 1).
Intercepts did not differ for small-leaved species (Table 1).

Disc-based estimates of individual mean LMA also showed positive correlations with whole-leaf estimates but their \emph{R\textsuperscript{2}} values were smaller than those for the species means (Fig. 2).
Similar to the species-level pattern, individuals with thicker and smaller leaves had stronger correlations (Fig. 2).
The relationships of the whole-leaf LMA : disc leaf LMA ratio with total dry mass for the leaf disc (we weighted 2-3 discs together), with leaf thickness, and with leaf area were heteroscedastic (unequal variance, \emph{P} \textless{} 0.001; Appendix S3).
Samples with small total dry mass, thin leaves, and larger leaves tended to show greater variance (Fig. 2; Appendix S3).
The total dry mass of leaf disc was positively correlated with both leaf thickness and leaf punch size (Appendix S4).
Amounts of intraspecific variation (CV; coefficient of variation) only weakly matched between whole-leaf and disc-based estimates (Fig. 3; \emph{R\textsuperscript{2}} = 0.33).

\hypertarget{discussion}{%
\section{DISCUSSION}\label{discussion}}

Disc-based traits explained interspecific variations in whole-leaf LMA (\emph{R\textsuperscript{2}} = , \emph{n} = ) quite well, and the strength of their relationship was approximately same or slightly weaker than previously reported (\emph{R\textsuperscript{2}} = 0.92, \emph{n} = 409, for whole-leaves in Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} = 0.92, \emph{n} = 364 for leaf laminas in Onoda et al. (\protect\hyperlink{ref-Onoda2011}{2011})).
Standardized major axis (SMA) regression analysis showed significantly greater intercepts when all the species were pooled (9.4\% differences) or large-leaved species were used (22.1\% differences).
Together, these findings suggest that disc-based estimates can be generally used as good proxies for their whole-leaf estimates with appropriate calibration.
However, the strength of correlations highly depended on leaf area, thickness (Fig. 1) and sizes of leaf punches (Appendix S3 and S5).

Larger leaves generally invest disproportionally more mass into dense midribs and major veins for support (\protect\hyperlink{ref-Niinemets2006}{Niinemets et al., 2006}, \protect\hyperlink{ref-Niinemets2007}{2007}; \protect\hyperlink{ref-Sack2012}{Sack et al., 2012}), and thinner leaves have distinct veins and have kite-type appearances (\protect\hyperlink{ref-Grubb1986}{Grubb, 1986}).
As we predicted, thicker and/or smaller leaves tended to show greater \emph{R\textsuperscript{2}} values between whole-leaf and leaf disc estimates (Figs 1. and 2).
Large-leaved species had greater whole-leaf LMA than leaf disc LMA (Table 1), which is consistent with the pattern of large leaves disproportionately investing in more leaf veins and/or petioles.
This is also supported by the subset of our dataset in which we measured leaf lamina and petiole dry mass separately.
We confirmed that leaf disc and whole-leaf LMA estimate diverged for species with greater investment in petiole dry mass, and large-leaved species tended to have greater investment for petiole dry mass than small-leaved species (Appendix S6).

Meanwhile, our results are seemingly inconsistent with a previous study that used thin and soft leaves and found high \emph{R\textsuperscript{2}} values (\emph{R\textsuperscript{2}} = 0.92; Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008})).
Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008}) used sapling leaves under closed canopy that are thinner than adult leaves (\protect\hyperlink{ref-Kitajima2010}{Kitajima and Poorter, 2010}; \protect\hyperlink{ref-Oktavia2020}{Oktavia and Jin, 2020}).
Shade leaves tend to have thinner leaves than sun leaves because of fewer layers of palisade mesophyll cells (\protect\hyperlink{ref-Terashima2001}{Terashima et al., 2001}; \protect\hyperlink{ref-Onoda2008}{Onoda et al., 2008}).
The relatively low \emph{R\textsuperscript{2}} values for thin-leaved species and individuals in our study compared to the previous study may suggest not only the effect of distinct veins on variance in LMA (\protect\hyperlink{ref-John2017}{John et al., 2017}) but also technical difficulty in measuring tiny weight of discs in lower LMA leaves.
We found that estimates of whole-leaf and leaf disc LMA diverged when total dry mass of leaf discs was small (Fig. 2; Appendices S3), and thus thin-leaved species, which have smaller dry mass (Appendix S4), tended to showed the low \emph{R\textsuperscript{2}} values especially when we used a smaller leaf punch (diameter: 0.6 cm; Appendix S5).

Even though we used the same species, the same individuals and the same trait (LMA) to estimate the extents of intraspecific trait variation (ITV), the explained variance of the relationship between whole-leaf and disc-based estimate was surprisingly small (Fig. 3; \emph{R\textsuperscript{2}} = 0.33).
This is probably because leaf disc samples have more sources of ITV than whole-leaf samples (i.e., variation among leaf discs within the same leaves).
Although there are not many studies that quantified trait variation within a single leaf, a previous study that used a giant leaved species \emph{Alocasia macrorrhiza}, suggests that LMA varied within leaves due to reduced water supply and demand from midrib to outer regions (\protect\hyperlink{ref-Li2013a}{Li et al., 2013}).
Because calculating the mean values generally diminishes the effect of within-sample variation, the relationship between disc-based estimates of individual mean LMA and its whole-leaf estimates will be stronger than the relationship between leaf disc CV and whole-leaf CV.
Thus, leaf disc and whole-leaf estimates of species means will show the strongest relationship.
This is consistent with the \emph{R\textsuperscript{2}} values in our results (Figs. 1-3).
The mismatch of ITV between whole-leaf and disc-based estimate could be ignored in community-level ITV studies (i.e., \protect\hyperlink{ref-Siefert2015}{Siefert et al., 2015}), which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies (i.e., \protect\hyperlink{ref-Albert2010a}{Albert et al., 2010}), which quantify extents of ITV for each species, since species with high ITV based on leaf-disc estimates do not always have to show high ITV for whole-leaf estimates.
This problem will be more obvious when one uses smaller leaf punches (Appendix S7).

The present study was designed to examine the extent to which whole leaf and leaf disc LMA matched.
Our study shows that mean values of leaf disc LMA is generally a good proxy for mean values of whole-leaf LMA.
However, larger and thinner leaves tended to show the large discrepancies between the leaf disc and whole-leaf estimates, and thus when one needs to mix leaf disc LMA and whole-leaf LMA in a single analysis (e.g., meta-analysis), appropriate calibrations are necessary.
Although our study was not designed to evaluate the effects of diameter sizes of leaf punches, the sizes of leaf punches are also the important sources of variance between whole-leaf and leaf disc LMA.
This is probably because small dry mass inflates measurement errors and small leaf disc size inflates trait variation within leaves.
Additionally, although the mean values of leaf disc and whole-leaf LMA matched well, quantifying trait variation using leaf discs should be considered with caution, because variation among leaf discs within the same leaves (including measurement errors) seems to be large.
Leaf discs, which can target leaf lamina avoiding leaf veins, are useful and essential for biochemical determinations such as measurements of Rubisco activity (\protect\hyperlink{ref-Parry2002}{Parry et al., 2002}), because the photosynthetic activity occurs most in lamina rather than in veins for C\textsubscript{3} plants (\protect\hyperlink{ref-Hibberd2002}{Hibberd and Quick, 2002}; \protect\hyperlink{ref-Gao2018}{Gao et al., 2018}).
Leaf disc LMA is also the preferred method for quantifying leaf lamina density, which avoid effects of veins.
However, if the objective is to use leaf disc LMA as an alternative for whole-leaf LMA, using an appropriate size of leaf punch is important for obtaining stable estimates of leaf disc LMA that matches well with whole-leaf LMA.

\hypertarget{acknowledgments}{%
\section{ACKNOWLEDGMENTS}\label{acknowledgments}}

The authors thank Lian-Bin Tao and Peng-Yun Yan for field work.
This work was funded by the National Natural Science Foundation of China (41861144016, 32071735, 31570406, 32171507), the CAS `Light of West China' Program, Yunnan Provincial Science and Technology Department (2018HB068).
PM was supported by financial support from CAS-TWAS President's Fellowship for International Doctoral Students.
MK was supported by CAS President's International Fellowship Initiative (2020FYB0003).

\hypertarget{author-contributions}{%
\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}}

PM, MK, and YJC conceived the study;
PM, YO, and YJC collected data;
MK and CZ performed the analysis;
MK and YJC led the writing of the paper; and
all authors contributed to revisions.

\hypertarget{data-availability-statement}{%
\section{DATA AVAILABILITY STATEMENT}\label{data-availability-statement}}

Data and codes for this manuscript are archived and available on Zenodo (XXX).

\hypertarget{supporting-information}{%
\section{SUPPORTING INFORMATION}\label{supporting-information}}

Additional supporting information may be found online in the Supporting Information section at the end of the article.

\textbf{APPENDIX S1.}
Site information.

\textbf{APPENDIX S2.}
Generalization for a relationship between ratios of disc-based and whole-leaf estimates of leaf mass per area (LMA), leaf tissue density (LD), and leaf thickness (LT).

\textbf{APPENDIX S3.}
Relationships between whole-leaf LMA : disc leaf LMA and (a) total dry mass for the leaf disc, (b) leaf thickness, and (c) leaf area.

\textbf{APPENDIX S4.}
Summary of multiple regression for log-transformed total dry mass of leaf discs.

\textbf{APPENDIX S5.}
Relationships between species mean leaf mass per area (LMA) determined by using whole leaves and leaf discs.

\textbf{APPENDIX S6.}
Relationships between petiole to leaf dry mass ratio and (a) leaf area and (b) whole-leaf to leaf disc LMA ratio.

\textbf{APPENDIX S7.}
Relationships between coefficient of variation (CV) in LMA determined by using whole leaves and leaf discs: (a) 0.6 cm diameter leaf punch, and (b) 1.0 cm diameter leaf punch.

\hypertarget{literature-cited}{%
\section{LITERATURE CITED}\label{literature-cited}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Albert2010a}{}}%
Albert, C. H., W. Thuiller, N. G. Yoccoz, A. Soudant, F. Boucher, P. Saccone, and S. Lavorel. 2010. \href{https://doi.org/10.1111/j.1365-2745.2010.01651.x}{Intraspecific functional variability: {Extent}, structure and sources of variation}. \emph{Journal of Ecology} 98: 604--613.

\leavevmode\vadjust pre{\hypertarget{ref-Carpenter2017}{}}%
Carpenter, B., A. Gelman, M. D. Hoffman, D. Lee, B. Goodrich, M. Betancourt, M. Brubaker, et al. 2017. \href{https://doi.org/10.18637/jss.v076.i01}{Stan : {A Probabilistic Programming Language}}. \emph{Journal of Statistical Software} 76: 1--32.

\leavevmode\vadjust pre{\hypertarget{ref-Diaz2004}{}}%
Díaz, S., J. G. Hodgson, K. Thompson, M. Cabido, J. H. C. Cornelissen, A. Jalili, G. Montserrat-Martí, et al. 2004. \href{https://doi.org/10.1658/1100-9233(2004)015\%5B0295:TPTTDE\%5D2.0.CO;2}{The plant traits that drive ecosystems: {Evidence} from three continents}. \emph{Journal of Vegetation Science} 15: 295.

\leavevmode\vadjust pre{\hypertarget{ref-Diaz2016}{}}%
Díaz, S., J. Kattge, J. H. C. Cornelissen, I. J. Wright, S. Lavorel, S. Dray, B. Reu, et al. 2016. \href{https://doi.org/10.1038/nature16489}{The global spectrum of plant form and function}. \emph{Nature} 529: 167--171.

\leavevmode\vadjust pre{\hypertarget{ref-Evans1972}{}}%
Evans, G. C. 1972. The quantitative analysis of plant growth. {Univ of California Press}.

\leavevmode\vadjust pre{\hypertarget{ref-Falster2018}{}}%
Falster, D. S., R. A. Duursma, and R. G. FitzJohn. 2018. \href{https://doi.org/10.1073/pnas.1714044115}{How functional traits influence plant growth and shade tolerance across the life cycle}. \emph{Proceedings of the National Academy of Sciences} 115: E6789--E6798.

\leavevmode\vadjust pre{\hypertarget{ref-Gao2018}{}}%
Gao, Z., W. Shen, and G. Chen. 2018. \href{https://doi.org/10.1093/jxb/ery155}{Uncovering {C4-like} photosynthesis in {C3} vascular cells}. \emph{Journal of Experimental Botany} 69: 3531--3540.

\leavevmode\vadjust pre{\hypertarget{ref-Garnier1994}{}}%
Garnier, E., and G. Laurent. 1994. \href{https://doi.org/10.1111/j.1469-8137.1994.tb04036.x}{Leaf anatomy, specific mass and water content in congeneric annual and perennial grass species}. \emph{New Phytologist} 128: 725--736.

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2013}{}}%
Gelman, A., J. B. Carlin, H. S. Stern, D. B. Dunson, A. Vehtari, and D. B. Rubin. 2013. Bayesian {Data Analysis}, {Third Edition}. {Taylor \& Francis}.

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008}{}}%
Gelman, A., A. Jakulin, M. G. Pittau, and Y. S. Su. 2008. \href{https://doi.org/10.1214/08-AOAS191}{A weakly informative default prior distribution for logistic and other regression models}. \emph{Annals of Applied Statistics} 2: 1360--1383.

\leavevmode\vadjust pre{\hypertarget{ref-Grubb1986}{}}%
Grubb, P. 1986. Sclerophylls, pachyphylls, and pycnophylls: {The} nature and significance of hard leaf surface. \emph{Insects and the Plant Surface}: 137--150.

\leavevmode\vadjust pre{\hypertarget{ref-Hanson1917}{}}%
Hanson, H. C. 1917. Leaf-structure as related to environment. \emph{American Journal of Botany}: 533--560.

\leavevmode\vadjust pre{\hypertarget{ref-Hibberd2002}{}}%
Hibberd, J. M., and W. P. Quick. 2002. \href{https://doi.org/10.1038/415451a}{Characteristics of {C4} photosynthesis in stems and petioles of {C3} flowering plants}. \emph{Nature} 415: 451--454.

\leavevmode\vadjust pre{\hypertarget{ref-John2017}{}}%
John, G. P., C. Scoffoni, T. N. Buckley, R. Villar, H. Poorter, and L. Sack. 2017. \href{https://doi.org/10.1111/ele.12739}{The anatomical and compositional basis of leaf mass per area}. \emph{Ecology Letters} 20: 412--425.

\leavevmode\vadjust pre{\hypertarget{ref-Katabuchi2015}{}}%
Katabuchi, M. 2015. \href{https://doi.org/10.1007/s11284-015-1307-x}{{LeafArea}: An {R} package for rapid digital image analysis of leaf area}. \emph{Ecological Research} 30: 1073--1077.

\leavevmode\vadjust pre{\hypertarget{ref-Kattge2020}{}}%
Kattge, J., G. Bönisch, S. Díaz, S. Lavorel, I. C. Prentice, P. Leadley, S. Tautenhahn, et al. 2020. \href{https://doi.org/10.1111/gcb.14904}{{TRY} plant trait database \textendash{} enhanced coverage and open access}. \emph{Global Change Biology} 26: 119--188.

\leavevmode\vadjust pre{\hypertarget{ref-Kichenin2013}{}}%
Kichenin, E., D. A. Wardle, D. A. Peltzer, C. W. Morse, and G. T. Freschet. 2013. \href{https://doi.org/10.1111/1365-2435.12116}{Contrasting effects of plant inter- and intraspecific variation on community-level trait measures along an environmental gradient}. \emph{Functional Ecology} 27: 1254--1261.

\leavevmode\vadjust pre{\hypertarget{ref-Kitajima2010}{}}%
Kitajima, K., and L. Poorter. 2010. \href{https://doi.org/10.1111/j.1469-8137.2010.03212.x}{Tissue-level leaf toughness, but not lamina thickness, predicts sapling leaf lifespan and shade tolerance of tropical tree species}. \emph{New Phytologist} 186: 708--721.

\leavevmode\vadjust pre{\hypertarget{ref-Kraft2008}{}}%
Kraft, N. J. B., R. Valencia, and D. D. Ackerly. 2008. \href{https://doi.org/10.1126/science.1160662}{Functional {Traits} and {Niche-Based Tree Community Assembly} in an {Amazonian Forest}}. \emph{Science} 322: 580--582.

\leavevmode\vadjust pre{\hypertarget{ref-Li2013a}{}}%
Li, S., Y.-J. Zhang, L. Sack, C. Scoffoni, A. Ishida, Y.-J. Chen, and K.-F. Cao. 2013. \href{https://doi.org/10.1371/journal.pone.0066016}{The heterogeneity and spatial patterning of structure and physiology across the leaf surface in giant leaves of alocasia macrorrhiza}. \emph{PLOS ONE} 8: 1--10.

\leavevmode\vadjust pre{\hypertarget{ref-Messier2017}{}}%
Messier, J., B. J. McGill, B. J. Enquist, and M. J. Lechowicz. 2017. \href{https://doi.org/10.1111/ecog.02006}{Trait variation and integration across scales: Is the leaf economic spectrum present at local scales?} \emph{Ecography} 40: 685--697.

\leavevmode\vadjust pre{\hypertarget{ref-Messier2010}{}}%
Messier, J., B. J. McGill, and M. J. Lechowicz. 2010. \href{https://doi.org/10.1111/j.1461-0248.2010.01476.x}{How do traits vary across ecological scales? {A} case for trait-based ecology}. \emph{Ecology Letters} 13: 838--848.

\leavevmode\vadjust pre{\hypertarget{ref-Niinemets2007}{}}%
Niinemets, Ü., A. Portsmuth, D. Tena, M. Tobias, S. Matesanz, and F. Valladares. 2007. \href{https://doi.org/10.1093/aob/mcm107}{Do we underestimate the importance of leaf size in plant economics? {Disproportional} scaling of support costs within the spectrum of leaf physiognomy}. \emph{Annals of Botany} 100: 283--303.

\leavevmode\vadjust pre{\hypertarget{ref-Niinemets2006}{}}%
Niinemets, Ü., A. Portsmuth, and M. Tobias. 2006. \href{https://doi.org/10.1111/j.1469-8137.2006.01741.x}{Leaf size modifies support biomass distribution among stems, petioles and mid-ribs in temperate plants}. \emph{New Phytologist} 171: 91--104.

\leavevmode\vadjust pre{\hypertarget{ref-Oktavia2020}{}}%
Oktavia, D., and G. Jin. 2020. \href{https://doi.org/10.1016/j.baae.2020.09.010}{Variations in leaf morphological and chemical traits in response to life stages, plant functional types, and habitat types in an old-growth temperate forest}. \emph{Basic and Applied Ecology} 49: 22--33.

\leavevmode\vadjust pre{\hypertarget{ref-Onoda2008}{}}%
Onoda, Y., F. Schieving, and N. P. R. Anten. 2008. \href{https://doi.org/10.1093/aob/mcn013}{Effects of light and nutrient availability on leaf mechanical properties of {Plantago} major: {A} conceptual approach}. \emph{Annals of Botany} 101: 727--736.

\leavevmode\vadjust pre{\hypertarget{ref-Onoda2011}{}}%
Onoda, Y., M. Westoby, P. B. Adler, A. M. F. Choong, F. J. Clissold, J. H. C. Cornelissen, S. Díaz, et al. 2011. \href{https://doi.org/10.1111/j.1461-0248.2010.01582.x}{Global patterns of leaf mechanical properties}. \emph{Ecology Letters} 14: 301--312.

\leavevmode\vadjust pre{\hypertarget{ref-Onoda2017}{}}%
Onoda, Y., I. J. Wright, J. R. Evans, K. Hikosaka, K. Kitajima, Ü. Niinemets, H. Poorter, et al. 2017. \href{https://doi.org/10.1111/nph.14496}{Physiological and structural tradeoffs underlying the leaf economics spectrum}. \emph{New Phytologist} 214: 1447--1463.

\leavevmode\vadjust pre{\hypertarget{ref-Parry2002}{}}%
Parry, M. A., P. J. Andralojc, S. Khan, P. J. Lea, and A. J. Keys. 2002. Rubisco activity: Effects of drought stress. \emph{Annals of Botany} 89: 833--839.

\leavevmode\vadjust pre{\hypertarget{ref-Perez-Harguindeguy2013}{}}%
Pérez-Harguindeguy, N., S. Díaz, E. Garnier, S. Lavorel, H. Poorter, P. Jaureguiberry, M. S. Bret-Harte, et al. 2013. \href{https://doi.org/10.1071/BT12225}{New handbook for standardised measurement of plant functional traits worldwide}. \emph{Australian Journal of Botany} 61: 167--234.

\leavevmode\vadjust pre{\hypertarget{ref-Poorter2014}{}}%
Poorter, H., H. Lambers, and J. R. Evans. 2014. \href{https://doi.org/10.1111/nph.12547}{Trait correlation networks: {A} whole-plant perspective on the recently criticized leaf economic spectrum}. \emph{New Phytologist} 201: 378--382.

\leavevmode\vadjust pre{\hypertarget{ref-Poorter2009}{}}%
Poorter, H., Ü. Niinemets, L. Poorter, I. J. Wright, and R. Villar. 2009. \href{https://doi.org/10.1111/j.1469-8137.2009.02830.x}{Causes and consequences of variation in leaf mass per area ({LMA}): {A} meta-analysis}. \emph{New Phytologist} 182: 565--588.

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
R Core Team. 2022. R: {A} language and environment for statistical computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.

\leavevmode\vadjust pre{\hypertarget{ref-Reich2014}{}}%
Reich, P. B. 2014. \href{https://doi.org/10.1111/1365-2745.12211}{The world-wide 'fast-slow' plant economics spectrum: {A} traits manifesto}. \emph{Journal of Ecology} 102: 275--301.

\leavevmode\vadjust pre{\hypertarget{ref-Sack2012}{}}%
Sack, L., C. Scoffoni, A. D. McKown, K. Frole, M. Rawls, J. C. Havran, H. Tran, and T. Tran. 2012. Developmentally based scaling of leaf venation architecture explains global ecological patterns. \emph{Nature Communications} 3: 1--10.

\leavevmode\vadjust pre{\hypertarget{ref-Siefert2015}{}}%
Siefert, A., C. Violle, L. Chalmandrier, C. H. Albert, A. Taudiere, A. Fajardo, L. W. Aarssen, et al. 2015. \href{https://doi.org/10.1111/ele.12508}{A global meta-analysis of the relative extent of intraspecific trait variation in plant communities}. \emph{Ecology Letters} 18: 1406--1419.

\leavevmode\vadjust pre{\hypertarget{ref-Terashima2001}{}}%
Terashima, I., S.-I. Miyazawa, and Y. T. Hanba. 2001. \href{https://doi.org/10.1007/PL00013972}{Why are {Sun Leaves Thicker} than {Shade Leaves}? \textemdash{} {Consideration} based on {Analyses} of {CO2 Diffusion} in the {Leaf}}. \emph{Journal of Plant Research} 114: 93--105.

\leavevmode\vadjust pre{\hypertarget{ref-Warton2012a}{}}%
Warton, D. I., R. A. Duursma, D. S. Falster, and S. Taskinen. 2012. \href{https://doi.org/10.1111/j.2041-210X.2011.00153.x}{Smatr 3\textendash{} an {R} package for estimation and inference about allometric lines}. \emph{Methods in Ecology and Evolution} 3: 257--259.

\leavevmode\vadjust pre{\hypertarget{ref-Warton2006}{}}%
Warton, D. I., I. J. Wright, D. S. Falster, and M. Westoby. 2006. \href{https://doi.org/10.1017/S1464793106007007}{Bivariate line-fitting methods for allometry}. \emph{Biological Reviews of the Cambridge Philosophical Society} 81: 259--291.

\leavevmode\vadjust pre{\hypertarget{ref-Wright2004a}{}}%
Wright, I. J., P. B. Reich, M. Westoby, D. D. Ackerly, Z. Baruch, F. Bongers, J. Cavender-Bares, et al. 2004. \href{https://doi.org/10.1038/nature02403}{The worldwide leaf economics spectrum}. \emph{Nature} 428: 821--827.

\leavevmode\vadjust pre{\hypertarget{ref-Wright2010}{}}%
Wright, S. J., K. Kitajima, N. J. B. Kraft, P. B. Reich, I. J. Wright, D. E. Bunker, R. Condit, et al. 2010. \href{https://doi.org/10.1890/09-2335.1}{Functional traits and the growth-mortality trade-off in tropical trees}. \emph{Ecology} 91: 3664--3674.

\leavevmode\vadjust pre{\hypertarget{ref-Yang2020}{}}%
Yang, J., J. Lu, Y. Chen, E. Yan, J. Hu, X. Wang, and G. Shen. 2020. Large {Underestimation} of {Intraspecific Trait Variation} and {Its Improvements}. \emph{Frontiers in Plant Science} 11.

\end{CSLReferences}

\newpage

\begin{longtable}[]{@{}lllr@{}}
\toprule()
Data & Slope & Intercept & \emph{R\textsuperscript{2}} \\
\midrule()
\endhead
All & 0.97 {[}0.94, 1.01{]} & \textbf{0.09 {[}0.02, 0.16{]}} & 0.87 \\
Thick-large & 0.93 {[}0.85, 1.03{]} & \textbf{0.19 {[}0.01, 0.37{]}} & 0.86 \\
Thick-small & 1.02 {[}0.97, 1.07{]} & 0 {[}-0.11, 0.11{]} & 0.93 \\
Thin-large & 0.91 {[}0.82, 1.02{]} & \textbf{0.2 {[}0.02, 0.37{]}} & 0.71 \\
Thin-small & 0.93 {[}0.83, 1.03{]} & 0.16 {[}-0.02, 0.35{]} & 0.84 \\
\bottomrule()
\end{longtable}

\textbf{Table. 1.}
Relationships between species mean leaf leaf mass per area (LMA) values determined by using whole leaves and leaf discs across all species and within species groups.
Data are divided into four categories based on the quartiles:
thin and small-leaved,
thin and large-leaved,
thick and small-leaved,
and thick and large-leaved species.
SMA slope (95\% CI), intercepts (95\% CI), and \emph{R\textsuperscript{2}} values are reported.
Statistically significant differences (\emph{P} \textless{} 0.05) in slopes and intercepts are highlighted in bold.

\newpage

\textbf{Fig. 1.}
Relationships between species mean leaf mass per area (LMA) determined by using whole leaves and leaf discs.
Species groups are divided into four categories on the basis of the medians of the leaf size and leaf thickness across all the species.
Dashed lines indicate 1:1 lines.
Blue solid lines indicate standardized major axis regressions.
The 95\% confidence intervals are presented as the shaded area.
All the correlations are significant (\emph{P} \textless{} 0.001).

\includegraphics{../figs/lalt_pool_grid.png}

\newpage

\textbf{Fig. 2.}
Relationships between individual leaf mass per area (LMA) determined by using whole leaves and leaf discs.
Individual groups are divided into four categories based on the basis of the medians of the leaf size and leaf thickness across all the individuals.
All the correlations are significant (\emph{P} \textless{} 0.001).
Details as in Fig. 1.

\includegraphics{../figs/lalt_tree_grid.png}

\newpage

\textbf{Fig. 3.}
Relationships between coefficient of variation (CV) in leaf trait values determined by using whole leaves and leaf discs: (a) leaf mass per area (LMA).
The correlation is significant (\emph{P} \textless{} 0.001).
Details as in Fig. 1.
Note that the axes are square root scales.

\includegraphics{../figs/cv_pool.png}

\end{document}