### 2022-05-07 report.Rmd
clean
- use format = "file"
- add site_info.json

---
title: Figures for "Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)"
author: "Masatoshi Katabuchi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
     collapsed: TRUE
     smooth_scroll: TRUE
     toc: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r}
library(tidyverse)
library(targets)
library(kableExtra)
```

```{r}
tar_load(sp_mean)
sp_mean |>
  group_by(ldlalt_gr) |>
  summarise(
    lma_leaf = mean(lma_leaf) |> round(1),
    lma_disc = mean(lma_disc) |> round(1)
  ) |>
  write_csv("data/ldlalt.csv")
```

# Leaf disc vs whole-leaf LMA (species-level)

## Pooled

<!-- I put tar_load to use {{targets}} -->

```{r, include=FALSE}
tar_load(lalt_pool_grid_png)
```

![](figs/lalt_pool_grid.png)

## Seprated

```{r, include=FALSE}
tar_load(lalt_sep_grid_png)
```

## MCMC check

```{r,eval=TRUE}

targets::tar_load(fit_sp_no_lma_summary_model)
targets::tar_load(fit_sp_no_lma_int_summary_model)
targets::tar_load(fit_sp_lma_summary_model)
targets::tar_load(fit_sp_lma2_summary_model2)
targets::tar_load(fit_sp_ld_summary_model)
targets::tar_load(fit_sp_ld2_summary_model2)
targets::tar_load(fit_sp_ld3_summary_model)
targets::tar_load(fit_sp_lm_summary_model)
targets::tar_load(fit_sp_lma0_summary_simple)

targets::tar_load(fit_sp_no_lma_diagnostics_model)
targets::tar_load(fit_sp_no_lma_int_diagnostics_model)
targets::tar_load(fit_sp_lma_diagnostics_model)
targets::tar_load(fit_sp_lma2_diagnostics_model2)
targets::tar_load(fit_sp_ld_diagnostics_model)
targets::tar_load(fit_sp_ld2_diagnostics_model2)
targets::tar_load(fit_sp_ld3_diagnostics_model)
targets::tar_load(fit_sp_lm_diagnostics_model)
targets::tar_load(fit_sp_lma0_diagnostics_simple)
```

```{r,eval=TRUE}
fit_sp_no_lma_summary_model |> filter(rhat > 1.1)
fit_sp_no_lma_int_summary_model |> filter(rhat > 1.1)
fit_sp_lma_summary_model |> filter(rhat > 1.1)
fit_sp_lma2_summary_model2 |> filter(rhat > 1.1)
fit_sp_ld_summary_model |> filter(rhat > 1.1)
fit_sp_ld2_summary_model2 |> filter(rhat > 1.1)
fit_sp_ld3_summary_model |> filter(rhat > 1.1)
fit_sp_lm_summary_model |> filter(rhat > 1.1)
fit_sp_lma0_summary_simple |> filter(rhat > 1.1)
```

It is difficult to model heteroskedasticity and measurement errors at the same time.

```{r,eval=TRUE}
div_check(fit_sp_no_lma_diagnostics_model)
div_check(fit_sp_no_lma_int_diagnostics_model)
div_check(fit_sp_lma_diagnostics_model)
div_check(fit_sp_lma2_diagnostics_model2)
div_check(fit_sp_ld_diagnostics_model)
div_check(fit_sp_ld2_diagnostics_model2)
div_check(fit_sp_ld3_diagnostics_model)
div_check(fit_sp_lm_diagnostics_model)
div_check(fit_sp_lma0_diagnostics_simple)
```

```{r}
targets::tar_load(loo_model)
loo::loo_compare(
  loo_model[[1]], loo_model[[2]], loo_model[[3]],
  loo_model[[4]], loo_model[[5]], loo_model[[6]],
  loo_model[[7]], loo_model[[8]], loo_model[[9]],
  loo_model[[10]], loo_model[[11]],
  loo_model[[12]], loo_model[[13]]
)
```

```{r}
tar_read(cv_sp)
```

## coef

![](figs/coef_sp_punch1_add.png)

# SMA table

## species

```{r, eval=TRUE}
tar_load(sma_sp_tab)
sma_sp_tab |>
  kable() |>
  kable_styling()
```
```{r}
sma_sp_tab |>
  write_csv("data/sma_sp_tab.csv")
```

## species LD

```{r, eval=TRUE}
tar_load(sma_sp_ld_tab)
sma_sp_ld_tab |>
  kable() |>
  kable_styling()
```

```{r}
sma_sp_ld_tab |>
  write_csv("data/sma_sp_ld_tab.csv")
```

## species simple

```{r, eval=TRUE}
tar_load(sma_sp_2_tab)
sma_sp_2_tab |>
  kable() |>
  kable_styling()
```

```{r}
sma_sp_2_tab |>
  write_csv("data/sma_sp_2_tab.csv")
```

## tree

```{r, eval=TRUE}
tar_load(sma_tree_tab)
sma_tree_tab |>
  kable() |>
  kable_styling()
```

## tree LD

```{r, eval=TRUE}
tar_load(sma_tree_ld_tab)
sma_tree_ld_tab |>
  kable() |>
  kable_styling()
```

# Validation

```{r}
tar_load(cv_sp)
cv_sp
```

```{r}
tar_load(cv_tree)
cv_tree
```

# Leaf disc vs whole-leaf LMA (individual-level)

```{r, include=FALSE}
tar_load(lalt_tree_grid_png)
```

![](figs/lalt_tree_grid.png)

# Divergence (species-level)

```{r, include=FALSE}
tar_load(ratio_png)
```

![](figs/ratio.png)

# LMA and LD (species-level)

```{r, include=FALSE}
tar_load(lma_ld_png)
```

![](figs/lma_ld.png)

# CV

```{r, include=FALSE}
tar_load(cv_pool_png)
```

![](figs/cv_pool.png)

```{r, include=FALSE}
tar_load(cv_sep_png)
```

![](figs/cv_sep.png)

# Leaf support cost (species-level)

```{r, include=FALSE}
tar_load(petiole_png)
```

![](figs/petiole.png)

# Computing Environment

```{r}
devtools::session_info()
```

```{r}
tar_visnetwork()
```