### 2022-05-11 ms/LMA_method.Rmd

---
geometry: margin=1in
link-citations: yes
csl: templates/american-journal-of-botany.csl
bibliography: leaf_disc.bib
fontsize: 12pt
#font-famliy: "ebgaramond"
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{lineno}
  - \linenumbers
output:
  html_document:
    toc: FALSE
#  pdf_document:
  bookdown::pdf_document2:
    keep_md: false
    fig_caption: yes
    keep_tex: yes
    toc: no
    number_sections: no
    template: templates/eisvogel2.tex
    #template: null
    pandoc_arg:
      - "--filter"
      - "pandoc-crossref"
  bookdown::word_document2:
    fig_width: 6
    fig_caption: yes
    number_sections: no
    toc: FALSE
    reference_docx: templates/rmd_style.docx
    pandoc_args:
    - "--filter"
    - "pandoc-crossref"
    - "--mathml"
---

```{r caching, include=FALSE}
library(knitr)
library(tidyverse)
library(methods)
source("../scripts/rmd_func.R")
basename <- "LMA_method"
opts_chunk$set(fig.path = paste("figs/", sep = ""))
opts_chunk$set(
  tidy = FALSE, warning = FALSE, message = FALSE,
  comment = NA, verbose = TRUE, echo = FALSE
)
# PDF-based figures
opts_chunk$set(dev = "pdf")

# https://github.com/haozhu233/kableExtra/issues/477
options(kableExtra.auto_format = FALSE)
library(kableExtra)
val <- yaml::yaml.load_file("../values.yml")
```

Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per are (LMA)

$$ $$

Phisamai Maenpuen^1,2,3^,
Masatoshi Katabuchi^1\*^,
Yusuke Onoda^4^,
Cong Zhou^1,2^,
Jiao-Lin Zhang^1,3\*^,
Ya-Jun Chen^1,3,5^

$$ $$

^1^ CAS Key Laboratory of Tropical Forest Ecology, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Menglun, Yunnan 666303, China

^2^ University of Chinese Academy of Sciences, Beijing 100049, China

^3^ Center of Plant Ecology, Core Botanical Gardens, Chinese Academy of Sciences, Yunnan 666303, China

^4^ Graduate School of Agriculture, Kyoto University, Kyoto 606-8502, Japan

^5^ Savanna Ecosystem Research Station, Xishuangbanna Tropical Botanical Garden, Chinese Academy of Sciences, Yuanjiang, Yunnan 6663300, China

$$ $$

**Corresponding Authors**:

Masatoshi Katabuchi

E-mail: mattocci27@gmail.com

Jiao-Lin Zhang

Tel: +86 691 871 3046;
Fax: +86 691 871 5070;
E-mail: zjl@xtbg.org.cn

$$ $$

Manuscript received _______; revision accepted _______.

**Running title**: Leaf disc and whole-leaf LMA

\newpage

# ABSTRACT

**PREMISE:**
Leaf mass per area (LMA), which is an important functional trait in leaf economic spectrum and plant growth analysis, is measured from leaf discs or whole leaves.
These differences in the measurement methods may lead to large differences in the estimates of LMA values.

**METHODS:**
We determined LMA using leaf discs and whole-leaf (including petiole) for `r val$all_sp` woody species from a wide range of biomes (tropics, subtropics, savanna, and temperate) to examine what extent whole-leaf and leaf disc LMA match, whether the differences in the estimates are associated with leaf size and thickness, and whether intraspecifc variation for each species match.

**RESULTS:**
Disc-based estimates of species and individual mean LMA matched well whole-leaf estimates.
Thin and large leaves had weaker correlations between leaf disc and whole-leaf estimates.
Variability of LMA estimates within species (i.e., amount of intraspecific variation) only weakly matched between whole-leaf and disc-based estimates (*R^2^* = `r val$cv_r2`).
We also found that sample with small total dry mass of leaf discs tended to show greater differences between leaf disc and whole-leaf LMA.

**CONCLUSIONS:**
Mean values of leaf disc LMA are generally good proxies for mean values of whole-leaf LMA with appropriate calibration (9.4% differences in our samples), but their accuracy depends on leaf size, leaf thickness, and sizes of leaf punch.
Therefore, using an appropriate size of leaf punch is important for obtaining stable estimates of leaf disc LMA that matches well with whole-leaf LMA.
Quantifying trait variation using leaf disc should be considered with caution.

**KEY WORDS:**
intraspecific variation,
leaf density,
leaf economic spectrum,
leaf heterogeneity,
leaf punch,
leaf size,
leaf thickness,
petiole,
specific leaf area,
within-leaf variation

# INTRODUCTION

The primary function of leaves is to return photosynthetic revenue on the investment that has been made in constructing the leaf [i.e., leaf economic spectrum, @Wright2004a; @Westoby2013].
Leaf mass per area (LMA), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area, which is a key trait in the leaf economic spectrum [@Wright2004a; @Poorter2009; @Onoda2017].
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover [@Wright2004a].
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns [@Garnier1994; @Wright2004a; @Reich2014; @Diaz2016].
The LMA is also frequently used in plant growth analysis [@Evans1972; @Poorter2014; @Falster2018] because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR $\times$  LMR $\times$ LMA^-1^).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species [i.e., 'soft' trait, @Diaz2004].
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago [@Hanson1917].
Actually, LMA has one of the best species coverage in leaf traits of the TRY plant data database (16,460 species) and was also the most often requested leaf trait (2,977 out of 7,330 requests), followed by leaf by leaf nitrogen (N) contents per leaf dry mass [12,238 species, 1,938 requests, @Kattge2020 [TRY version 5, status October 1 2019]]

LMA values can be determined by either measuring whole-leaf (including or excluding petioles) or leaf disc samples [@Perez-Harguindeguy2013].
According to the TRY database version 5, recorded numbers of observations and species are as follows:
1) a whole leaf including (88,490 observations and 7,068 species) or excluding petioles (64,838 observations and 7,558 species; note that this could include leaf disc samples), and
2) a leaf disc excluding all major veins or petioles (645 observations and 403 species) [@Kattge2020 [TRY version 5]].
Although many studies do not make clear what protocol was followed (146,315 observations and 13,101 species), these differences in the measurement methods may lead to large discrepancies in the estimates of LMA values because the major vein allocation (major vein, which includes first to second or third-order veins, volume per area) has been reported to be one of the main determinants for the variation in LMA [@John2017].
Previous works that compared estimates of whole-leaf LMA and leaf disc LMA using tropical tree species showed good correlations between the two [*R^2^* = 0.92 for whole-leaves in @Kraft2008; *R^2^* = 0.92 for leaf laminas in @Onoda2011], but the comparisons between the two estimates in other biomes have rarely been examined.

Since larger leaves tend to invest more of their mass into dense midribs and petiole for support [@Niinemets2006; @Niinemets2007; @Li2022; @Li2022a], and thinner leaves have clearly visible and large-diameter veins with less uniform leaf structure (i.e., kite-type leaves; [@Grubb1986]), discrepancies in the estimates of whole-leaf LMA and leaf disc LMA might be greater for larger and thinner leaves.
(**add leaf density**)
If leaf disc LMA largely and consistently underestimates whole-leaf LMA (i.e., differences in means), some calibrations may be required to combine or compare those different estimates [@Kraft2008; @Onoda2011].
If divergencies between the estimates based on the different methods are large and inconsistent (i.e., differences in variances and low *R^2^* values), this is difficult to calibrate and should inflate errors in the subsequent analyses.
To date, however, only a database from a single region, Panama Plant Traits Database [@Wright2010], is available in the TRY [@Kattge2020] that have both estimates of LMA from small leaf discs and whole leaves including petioles, limiting of our understanding of under what conditions estimates of LMA from leaf discs are reliable estimates of whole-leaf LMA.

Intraspecific trait variation (ITV), which reflects both heritable genetic variation and phenotypic plasticity, influences plant responses to abiotic and biotic interactions (REF).
The extent of ITV within and among communities for LMA can be similar to or greater than interspecific trait variation [e.g., @Messier2010; @Kichenin2013; @Fajardo2018].
The coefficient of variation (CV) is often used for quantifying the extent of ITV for each species [@Yang2020; @Westerband2021].
Sources of ITV for whole-leaf samples are variation among individuals within the same species and variation among leaves within the same individuals [@Messier2010; @Messier2017].
An additional source of trait variation for leaf disc samples is variation among leaf discs within the same leaves.
Given that leaf disc samples have more sources of ITV than whole-leaf samples, the effects of discrepancies between whole-leaf LMA and leaf disc LMA might be large when ITV is quantified based on CV.
Despite the importance of ITV in community ecology [@Siefert2015; @Westerband2021], the effect of different measurement methods on the extent of ITV has been largely ignored.

In this study, we aimed to investigate the relationship between whole-leaf (including petiole) LMA and leaf disc LMA,
whether the relationship varied with leaf tissue density, leaf area and leaf thickness, and
whether the extent of ITV for each species match between whole-leaf and leaf disc based estimates.
We collected a total of `r val$all_sp` woody species from four biomes (tropics, subtropics, savanna, and warm-temperate) to cover the wide range of geography.
We predicted that whole-leaf LMA would be greater than leaf disc LMA because whole-leaf LMA includes petioles and midribs that have greater dry mass per unit area than laminas [@Niinemets2007].

We also predicted that species with larger and/or thinner leaves would show weaker correlations and larger differences in LMA estimates.
(**add leaf density**)

Finally, we predicted that the extent of ITV do not match well between whole-leaf and leaf disc based estimates because leaf disc samples have more sources of variation than whole-leaf samples.

# MATERIALS AND METHODS

*Data sources*

We used newly compiled individual-level plant datasets form a wide ranges of biomes in China and Japan (Table S1) to examine the relationship between whole-leaf and leaf disc-based estimates of leaf traits.
First, the Yunnan dataset is from three forest plots in Yunnan province, Southwest China, which includes a tropical rain forest (TRF), a tropical hot-dry savanna ecosystem (HDS), and a subtropical evergreen wet forest (STF), with a total of `r val$sp1` species and `r val$tree1` individuals.
Second, the Yakushima data is from a warm-temperate forest in Yakushima island, Japan, with a total of `r val$yaku_sp` species and `r val$tree3` individuals.
In total, our data set comprised `r val$all_sp` woody species and `r val$all_tree` individual that ranged from 8.1 to 24.7 °C in mean annual temperature (MAT) and 732.8 to 4477 mm in mean annual precipitation (MAP) [Table S1; @Eguchi2006; @Song2017; @Fei2018], which covers the wide climate ranges in the distribution of broad-leaved tree species.
The methodologies for trait sampling and measurement are slightly different between the Yunnan and the Yakushima dataset, which we describe bellow.

*Measurements of leaf disc and whole-leaf LMA*

At the TRF site, we collected leaf samples from trees within reach of a canopy crane (88 m tall with a 60 m long boom).
In the HDS, and STF sites, we used a 12-m long pruner to collect samples for most of the target species.
In case of tall individuals that out of the reach of pruner, we used a rope climbing to reach the canopy, then sampled branches with the long pruner.
In Yakushima, we used 15-m poles for our leaf sampling.
At least six sun-exposed healthy leaves were sampled from each of 3 to 6 individuals for each species.
Trait values were averaged at the species-level for the analyses.
For compound-leaved species, we referred a leaflet (the minimum photosynthetic unit) as a single leaf in our study.
We excluded species that had tiny leaves or leaflets (length <1 cm) because it was not practical to measure leaf disc LMA.

To determine what extent LMA values were affected by the different methods, LMA were determined (i) LMA based on a whole leaf including petioles and (ii) LMA based on leaf disc avoiding thick veins.
Three 0.6-cm diameter discs were taken from the base, middle and tip of each leaf using a hole punch in the Yunnan dataset.
Two 1.0-cm diameter discs were taken from each leaf were in the Yakushima dataset (see Fig. S1 for more details).
Fresh leaf area (LA; cm^-2^) of leaf materials (whole leaves including petioles and midribs) was measured using a scanner and ImageJ software by the R package *LeafArea* [@Katabuchi2015].
Fresh leaf thickness (LT; mm) was measured using a micrometer in three points at the base, middle and tip of the leaf using a micrometer (Mitutoyo 293-240, Mitutoyo Corporation, Kawasaki, Japan) with a precision at 0.001 mm.
Thickness of each leaf disc was also measured for the Yunnan dataset but not for the Yakushima dataset, which enables us to estimate whole-leaf and leaf disc leaf tissue density (LD; g cm^-3^) separately in the Yunnan dataset.
Leaf dry mass (whole leaves including petioles and midribs, and leaf discs) was recorded after drying 80°C for over 48 h to constant weight using an electronic scale (Mettler-Toledo MS204TS, Columbus, OH, U.S.A.) with a precision at 0.0001 g.
In the subset of Yakushima dataset, leaf lamina and petiole dry mass were also determined separately (`r val$pet_sp` out of `r val$yaku_sp` species) to quantify leaf support costs.
Dry mass for leaf discs were recorded 2-3 discs together for each leaf.
Based on those measurements of whole-leaf and leaf disc samples, LMA (g m^-2^) was calculated as the ratio of leaf dry mass to fresh leaf area, and LD was calculated as the ratio of LMA to LT.
We do not present results of LD in the main text, because differences between whole-leaf LD and leaf disc LD only depends on the ratio between whole-leaf LMA and leaf disc LMA and measurement errors of leaf thickness (Appendix S2).

*Relationship between whole-leaf and leaf disc LMA*

A bivariate trait relationship between disc-based and whole-leaf log-transformed LMA was summarized with variance explained (*R^2^*) based on Pearson correlations and with a standardized major axis (SMA) regression [@Warton2006] across all the species.

A SMA analysis was run using the R package *smatr* [@Warton2012a]; this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 2000 bootstraps.

*Effects of leaf morphology on the relationship between whole-leaf and leaf disc LMA*

To further quantify the effect of leaf morphology and leaf punch size on the relationship between whole-leaf and leaf disc LMA, we built a hierarchical Bayesian model.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of  leaf tissue density, leaf area and leaf thickness on the log-scale;

\begin{align}
\mathrm{ln}(LMA_{Wi}) & \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i^2) (\#eq:LMA)\\
\tilde{\mu_i} & = \beta_0 + \beta_1 \mathrm{ln}(LD_{i}) + \beta_2 \mathrm{ln}(LA_{i}) + \beta_3 \mathrm{ln}(LT_{i}) + \beta_4 PS_{i} \nonumber \\
& + PS_{i} \times \bigl\{ \beta_5 \mathrm{ln}(LD_{i}) + \beta_6 \mathrm{ln}(LA_{i}) + \beta_7 \mathrm{ln}(LT_{i})\bigr\}(\#eq:mu)
\end{align}

where $LMA_{Wi}$, $LMA_{Di}$, $LD_{i}$, $LA_{i}$, and $LT_{i}$ are the observed whole-leaf LMA, leaf disc LMA, leaf tissue density, leaf area, and leaf thickness for species *i*, respectively,
The dummy variable $PS_i$ is set to 1 for samples obtained with a small leaf punch (0.6-cm diameter) and 0 for samples obtained with a large leaf punch (1.0-cm diameter).
$\tilde{\mu_i}$ is the effect of leaf tissue density, leaf area, leaf thickness and leaf punch size on the mean estimates for the whole-leaf LMA of species *i*, $\sigma_i^2$ is the variance of whole-leaf LMA for species *i*, and $\beta$ are coefficients.
The standard deviation of whole-leaf LMA ($\sigma_i$) was assumed to follow a normal distribution;

\begin{align}
\mathrm{ln}(\sigma_i) & \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega^2) (\#eq:ln-sigma) \\
\mathrm{ln}(\tilde{\sigma_i}) & = \gamma_0 + \gamma_1 \mathrm{ln}(LD_{i}) + \gamma_2 \mathrm{ln}(LA_{i}) + \gamma_3 \mathrm{ln}(LT_{i}) + \gamma_4 PS_{i} \nonumber \\
& + PS_{i} \times \bigl\{\gamma_5 \mathrm{ln}(LD_{i}) + \gamma_6 \mathrm{ln}(LA_{i}) + \gamma_7 \mathrm{ln}(LT_{i})\bigr\}(\#eq:tilde-sigma)
\end{align}

where $\tilde{\sigma_i}$ is a hyperparameter for $\sigma_i$ ,
$\omega^2$ is a scaling hyperparameter, and $\gamma$ are coefficients.
To allow comparisons among parameter estimates, $\mathrm{ln}(LD)$, $\mathrm{ln}(LA)$, and $\mathrm{ln}(LT)$ were scaled to a mean of 0 and standard deviation of 1.
Note that this model does not include measurement errors of leaf disc LMA implemented in the previous SMA analysis, because there are too many scaling parameters

Posterior distributions of all parameters were estimated using the Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan [@Carpenter2017] using the weakly-informative priors [@Gelman2008].
See Appendix SX for more detail.
The Stan code use to fit models is available from Github at: [https://github.com/mattocci27/leaf-disc](https://github.com/mattocci27/leaf-disc).
Convergence of the posterior distribution was assessed with the Gelman-Rubin statistic with a convergence threshold of 1.1 for all parameters [@Gelman2013].

We also perfomred ordinary least-squares (OLS) regressions to see the relationship between whole-leaf to leaf disc LMA ratio and petiole to leaf dry mass ratio and between leaf area petiole to leaf dry mass ratio, using a subset of the dataset in which we measured leaf lamina and petiole dry mass separately.

*Relationship between disc-based and whole-leaf estiamtes of intraspecific variation*

The extent of intraspecific variation (ITV) was calculated using Bao's CV estimator [@Yang2020];

\begin{align}
CV = CV_1 - \frac{CV_1^3}{N} + \frac{CV_1}{4N} + \frac{CV_1^2 \gamma_1}{2N} + \frac{CV_1 \gamma_2}{8N} (\#eq:CV)
\end{align}

where $CV_1 = s/\bar{x}$, $s$ is the sample standard deviation and $\bar{x}$ is the sample mean, *N* is the sample size, and $\gamma_1$ and $\gamma_2$ are the skewness and kurtosis of the trait distribution.
Bao's CV estimator is a robust method to quantify ITV based on a coefficient of variation (CV) when sample size is small [@Yang2020].
We calulcated CV for species that have at least 5 indivdiuals as recommended in @Westerband2021.
Ideally, 20 individuals for each species are required for robust estimation of CV [i.e., <5% differences from the true ITV; @Yang2020].

SMA analyses were run as described above.
We also perfmoed the paired t-test to compare the extent of ITV between whole-leaf LMA and leaf disc LMA.
All statistical analyses were conducted in R version 4.1.3 [@RCoreTeam2022].

# RESULTS

There were a significant shift in elevation between whole-leaf and leaf disc LMA (Fig.1, SMA slope: `r val$sma_slope_all_mean` [95%CI: `r val$sma_slope_all_lwr`, `r val$sma_slope_all_upr`], SMA intercept: `r val$sma_int_all_mean` [95%CI: `r val$sma_int_all_lwr`, `r val$sma_int_all_upr`]), suggesting that the whole-leaf estimates of LMA were `r val$sma_per_all_mean`% [95%CI: `r val$sma_per_all_lwr`, `r val$sma_per_all_upr`] greater than disc-based estimates in the arithmetic scale.

Leaf tissue density, leaf area, leaf thickness and leaf punch size influenced the relationship between the whole-leaf LMA and leaf disc LMA (Fig. 2 and Appendix SX).
Species with lower leaf tissue density had greater variance in the estiamte of whole-leaf to leaf disc LMA ratio (Fig. 2a).
When a small leaf punch was used, this pattern was more pronounced and leaf disc and whole-leaf LMA estimate diverged for species with higher leaf tissue density (Fig. 2a).
Species with larger leaves had greater estimate of the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2b), which was only significant for leaves obtained by a large punch.
Species with thinner leaves had grater variance in the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2c), which was only significant for leaves obtained by a small punch.
A small punch also inflated variance in the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2a-c and Appendix SX).
Leaf disc and whole-leaf LMA estimate diverged for species with greater investment in petiole dry mass (Fig. 3a). Large-leaved species tended to have greater investment for petiole dry mass than small-leaved species (Fig. 3b).

The extent of intraspecific trait variation (CV; coefficient of variation) only weakly matched between whole-leaf and disc-based estimates (Fig.4).
Altough there were no stiatical differences in the SMA slope or elevation (SMA slope: `r val$sma_slope_cv_mean` [95%CI: `r val$sma_slope_cv_lwr`, `r val$sma_slope_cv_upr`], SMA intercept: `r val$sma_int_cv_mean` [95%CI: `r val$sma_int_cv_lwr`, `r val$sma_int_cv_upr`]), leaf disc LMA had greater CVs than whole-leaf LMA (*t* = `r val$t_test_t`, *d.f.* = `r val$t_test_df`, *P* < 0.001).

# DISCUSSION

Disc-based traits explained interspecific variations in whole-leaf LMA (*R^2^* = `r val$r2_LMA`, *n* = `r val$N3`) quite well, and the strength of their relationship was approximately same or slightly weaker than previously reported  (*R^2^* = 0.92, *n* = 409,  for whole-leaves in @Kraft2008, *R^2^* = 0.92, *n* = 364 for leaf laminas in @Onoda2011).
Standardized major axis (SMA) regression analysis showed significantly greater intercepts when all the species were pooled (9.4% differences) or large-leaved species were used (22.1% differences).
Together, these findings suggest that disc-based estimates can be generally used as good proxies for their whole-leaf estimates with appropriate calibration.
However, the strength of correlations highly depended on leaf area, thickness (Fig. 1) and sizes of leaf punches (Appendix S3 and S5).

Larger leaves generally invest disproportionally more mass into dense midribs and major veins for support [@Niinemets2006; @Niinemets2007; @Sack2012], and thinner leaves have distinct veins and have kite-type appearances [@Grubb1986].
As we predicted, thicker and/or smaller leaves tended to show greater *R^2^* values between whole-leaf and leaf disc estimates (Figs 1. and 2).
Large-leaved species had greater whole-leaf LMA than leaf disc LMA (Table 1), which is consistent with the pattern of large leaves disproportionately investing in more leaf veins and/or petioles.
This is also supported by the subset of our dataset in which we measured leaf lamina and petiole dry mass separately.
We confirmed that leaf disc and whole-leaf LMA estimate diverged for species with greater investment in petiole dry mass, and large-leaved species tended to have greater investment for petiole dry mass than small-leaved species (Appendix S6).

Meanwhile, our results are seemingly inconsistent with a previous study that used thin and soft leaves and found high *R^2^* values (*R^2^* = 0.92; @Kraft2008).
@Kraft2008 used sapling leaves under closed canopy that are thinner than adult leaves [@Kitajima2010; @Oktavia2020].
Shade leaves tend to have thinner leaves than sun leaves because of fewer layers of palisade mesophyll cells [@Onoda2008; @Terashima2001].
The relatively low *R^2^* values for thin-leaved species and individuals in our study compared to the previous study may suggest not only the effect of distinct veins on variance in LMA [@John2017] but also technical difficulty in measuring tiny weight of discs in lower LMA leaves.
We found that estimates of whole-leaf and leaf disc LMA diverged when total dry mass of leaf discs was small (Fig. 2; Appendices S3), and thus thin-leaved species, which have smaller dry mass (Appendix S4), tended to showed the low *R^2^* values especially when we used a smaller leaf punch (diameter: 0.6 cm; Appendix S5).

Even though we used the same species, the same individuals and the same trait (LMA) to estimate the extents of intraspecific trait variation (ITV), the explained variance of the relationship between whole-leaf and disc-based estimate was surprisingly small (Fig. 3; *R^2^* = `r val$cv_r2`).
This is probably because leaf disc samples have more sources of ITV than whole-leaf samples (i.e., variation among leaf discs within the same leaves).
Although there are not many studies that quantified trait variation within a single leaf, a previous study that used a giant leaved species *Alocasia macrorrhiza*, suggests that LMA varied within leaves due to reduced water supply and demand from midrib to outer regions [@Li2013a].
Because calculating the mean values generally diminishes the effect of within-sample variation, the relationship between disc-based estimates of individual mean LMA and its whole-leaf estimates will be stronger than the relationship between leaf disc CV and whole-leaf CV.
Thus, leaf disc and whole-leaf estimates of species means will show the strongest relationship.
This is consistent with the *R^2^* values in our results (Figs. 1-3).
The mismatch of ITV between whole-leaf and disc-based estimate could be ignored in community-level ITV studies [i.e., @Siefert2015], which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies [i.e., @Albert2010a], which quantify extents of ITV for each species, since species with high ITV based on leaf-disc estimates do not always have to show high ITV for whole-leaf estimates.
This problem will be more obvious when one uses smaller leaf punches (Appendix S7).

The present study was designed to examine the extent to which whole leaf and leaf disc LMA matched.
Our study shows that mean values of leaf disc LMA is generally a good proxy for mean values of whole-leaf LMA.
However, larger and thinner leaves tended to show the large discrepancies between the leaf disc and whole-leaf estimates, and thus when one needs to mix leaf disc LMA and whole-leaf LMA in a single analysis (e.g., meta-analysis), appropriate calibrations are necessary.
Although our study was not designed to evaluate the effects of diameter sizes of leaf punches, the sizes of leaf punches are also the important sources of variance between whole-leaf and leaf disc LMA.
This is probably because small dry mass inflates measurement errors and small leaf disc size inflates trait variation within leave.
Additionally, although the mean values of leaf disc and whole-leaf LMA matched well, quantifying trait variation using leaf discs should be considered with caution, because variation among leaf discs within the same leaves (including measurement errors) seems to be large.
Leaf discs, which can target leaf lamina avoiding leaf veins, are useful and essential for biochemical determinations such as measurements of Rubisco activity [@Parry2002], because the photosynthetic activity occurs most in lamina rather than in veins for C~3~ plants [@Hibberd2002; @Gao2018].
Leaf disc LMA is also the preferred method for quantifying leaf lamina density, which avoid effects of veins.
However, if the objective is to use leaf disc LMA as an alternative for whole-leaf LMA, using an appropriate size of leaf punch is important for obtaining stable estimates of leaf disc LMA that matches well with whole-leaf LMA.

# ACKNOWLEDGMENTS

The authors thank Lian-Bin Tao and Peng-Yun Yan for field work.
This work was funded by the National Natural Science Foundation of China (41861144016, 32071735, 31570406, 32171507), the CAS 'Light of West China' Program, Yunnan Provincial Science and Technology Department (2018HB068).
PM was supported by financial support from CAS-TWAS President’s Fellowship for International Doctoral Students.
MK was supported by CAS President’s International Fellowship Initiative (2020FYB0003).

# AUTHOR CONTRIBUTIONS

PM, MK, and YJC conceived the study;
PM, YO, and YJC collected data;
MK and CZ performed the analysis;
MK and YJC led the writing of the paper; and
all authors contributed to revisions.

# DATA AVAILABILITY STATEMENT

Data and codes for this manuscript are archived and available on Zenodo (XXX).

# SUPPORTING INFORMATION

Additional supporting information may be found online in the Supporting Information section at the end of the article.

**APPENDIX S1.**
Site information.

**APPENDIX S2.**
Generalization for a relationship between ratios of disc-based and whole-leaf estimates of leaf mass per area (LMA), leaf tissue density (LD), and leaf thickness (LT).

**APPENDIX S3.**
Relationships between whole-leaf LMA : disc leaf LMA and (a) total dry mass for the leaf disc, (b) leaf thickness, and (c) leaf area.

**APPENDIX S4.**
Summary of multiple regression for log-transformed total dry mass of leaf discs.

**APPENDIX S5.**
Relationships between species mean leaf mass per area (LMA) determined by using whole leaves and leaf discs.

**APPENDIX S6.**
Relationships between petiole to leaf dry mass ratio and (a) leaf area and (b) whole-leaf to leaf disc LMA ratio.

**APPENDIX S7.**
Relationships between coefficient of variation (CV) in LMA determined by using whole leaves and leaf discs: (a) 0.6 cm diameter leaf punch, and (b) 1.0 cm diameter leaf punch.

# LITERATURE CITED

<div id="refs"></div>

\newpage

```{r tb1, echo=FALSE, eval=FALSE}
tab1 <- read.csv("../data/sma_sp_ld_tab.csv")
colnames(tab1) <- c("Data", "Slope", "Intercept", "*R^2^*")

tab1 <- tab1 |>
  mutate(Data = case_when(
    Data == "Dense~Thick~Large" ~ "Dense-thick-large",
    Data == "Dense~Thick~Small" ~ "Dense-thick-small",
    Data == "Dense~Thin~Large" ~ "Dense-thin-large",
    Data == "Dense~Thin~Small" ~ "Dense-thin-small",
    Data == "Nondense~Thick~Large" ~ "Nondense-thick-large",
    Data == "Nondense~Thick~Small" ~ "Nondense-thick-small",
    Data == "Nondense~Thin~Large" ~ "Nondense-thin-large",
    Data == "Nondense~Thin~Small" ~ "Nondense-thin-small",
    TRUE ~ Data
  ))
if ("html" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  tab1 |>
    kable(format = "html") |>
    kable_classic()
} else if ("docx" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  tab1 |>
    kable(format = "pipe")
} else if ("latex" == knitr::opts_knit$get("rmarkdown.pandoc.to")) {
  tab1 |>
    kable(format = "pipe")
}
```

**Fig. 1.**
Relationships between species mean leaf mass per area (LMA) determined by using whole-leaves and leaf discs.
Dashed lines indicate 1:1 lines.
Blue solid lines indicate standardized major axis (SMA) regressions.
The 95% confidence intervals are resented as the shaded area.
The correlation is significant (*P* < 0.001).

![](../figs/sma.png)

\newpage

**Fig. 2.**
Model predictions of the whole-leaf to leaf disc LMA ratio as a function of leaf punch sizes and (a) leaf tissue density, (b) leaf area, and (c) leaf thickness.
Dashed line indicates 1:1 ratio of leaf disc and whole-leaf estimates.
Solid lines indicate the posterior means ($\tilde{\mu}$ in Eq. 1-2), and the shade regions shows $\pm$ the posterior means of standard deviations ($\tilde{\sigma}$ in Eq. 3-4).

![](../figs/pred_mcmc.png)

\newpage

**Fig. 3.**
Relationships between petiole to leaf dry mass ratio and (a) leaf area and (b) whole-leaf to leaf disc LMA ratio.
Blue solid lines indicate ordinary least squares (OLS) regressions.
The 95% confidence intervals are presented as the shaded area.
All the correlations are significant (P < 0.001).
All the samples were obtained with a diameter 1.0-cm leaf punch.

![](../figs/petiole.png)

\newpage

**Fig. 4.**
Relationships between coefficient of variation (Bao's CV estimator, Eq. 5) in LMA values determined by using whole leaves and leaf discs.
The correlation is significant (*P* < 0.001).
Details as in Fig. 1.

![](../figs/cv_pool.png)