### 2022-08-09 ms/s4.Rmd

---
#title: "Appendix S1: Supporting Tables and Figures."
geometry: margin=1in
link-citations: yes
csl: templates/american-journal-of-botany.csl
bibliography: leaf_disc.bib
fontsize: 12pt
header-includes:
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{colortbl}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \usepackage[default]{sourcesanspro}
  - \usepackage{sourcecodepro}
  - \fancypagestyle{plain}{\pagestyle{fancy}}
  - \fancyhead[RE,RO]{Maenpuen \textit{et al}. --American Journal of Botany-- Appendix S4}
output:
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: no
    keep_tex: yes
---

# Appendix S4:

Standardized regression coefficients modeling the effects of leaf tissue density,
leaf area,
leaf thickness,
punch size, and
their interactions on
(a) the mean estimates of whole-leaf-based leaf mass per area (LMA) and
(b) the estimated variance of whole-leaf LMA.
The leaf disc LMA was used for the baseline mean of the whole-leaf LMA.
Thick and thin lines indicate 90% and 95% credible intervals, respectively.
Circles show posterior means of coefficients.
Circles filled with blue indicate significant effects and white indicate non-significance effects.

![](../figs/coef_sp_punch1_add.pdf)